### 2022-05-18 main/chapter_2ljets.tex
rjr former excesses and conventional comparison

\begin{figure}[tp]
\centering
\begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figures/2ljets_sr_high_8_met_sig.pdf}
    \caption{SR-High-8}
\end{subfigure}
\hfill
\begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figures/2ljets_sr_llbb_mbb.pdf}
    \caption{SR-$\ell\ell bb$}
\end{subfigure}
\\[0.5em]
\begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figures/2ljets_sr_int_met_sig.pdf}
    \caption{SR-Int}
\end{subfigure}
\hfill
\begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figures/2ljets_sr_low_met_sig.pdf}
    \caption{SR-Low}
\end{subfigure}
\caption{%
Example signal region histograms with benchmark signal sample yields overlaid
as dotted lines~\cite{atlas2022searches}.
Physically, signal yields would add on top of the backgrounds.
Data are shown; regions in the the top two plots observe zero data, and the
Poisson error bars for $0$ events are hidden for aesthetic reasons.
}
\label{fig:2ljets_signal_examples}
\end{figure}

@@

\begin{figure}[tp]
\centering
\includegraphics[width=0.99\textwidth]{figures/2ljets_contours_c1n2.pdf}
\caption{%
Contours for the C1N2 model in the $\twoljets$ electroweak
analysis~\cite{atlas2022searches}.
Space below the solid red line is labelled as excluded and its dotted
neighbours show the result if all signal cross-sections are varied up and down
by theoretical error bars.
The yellow band shows the $\pm1$-sigma region of exclusion contours
from asymptotic approximations to the prior distribution of the test statistic.
Grey areas are observed limits from the two-lepton parts
of~\cite{atlas_23l_SUSY_2016_24} and~\cite{atlas_2l_SUSY_2013_11}.
Exclusion is defined by the $95\%$ $\mathrm{CLs}$ prescription
in asymptotic approximations.
All contours are interpolated from a sparse grid.
}
\label{fig:2ljets_contours_c1n2}
\end{figure}

\begin{figure}[tp]
\centering
\includegraphics[width=0.99\textwidth]{figures/2ljets_contours_gmsb.pdf}
\caption{%
Contours for the GMSB model in the $\twoljets$ electroweak
analysis~\cite{atlas2022searches}.
Space below the solid red line is labelled as excluded and its dotted
neighbours show the result if all signal cross-sections are varied up and down
by theoretical error bars.
The yellow band shows the $\pm1$-sigma region of exclusion contours
from asymptotic approximations to the prior distribution of the test statistic.
Exclusion is defined by the $95\%$ $\mathrm{CLs}$ prescription
in asymptotic approximations.
All contours are interpolated from a sparse grid.
}
\label{fig:2ljets_contours_gmsb}
\end{figure}

@@

\clearpage
\FloatBarrier
\section{Origins}
% previous results on 2(/3)L from ATLAS
% other constraints on these models from ATLAS/CMS
% previous region design, motivation for this work
Previous analyses of \atlas\ data have explored similar selections of
two leptons, jets, and missing transverse momentum.
Quite a few analyses, in fact.
This project exists to supplement those results with information from the full
LHC Run 2 data-set which was completed in 2018 and contains
$139~\mathrm{fb}^{-1}$ of usable proton-proton collisions at a collision energy
of $13~\eV[T]$.
Leverage of the increased data quantity is one key aim.
Analysis quality also might be improved if our designs can take lessons from
the experiences of previous work.

@@

In this way, RJR variables attain intelligence in how parent particles are
reconstructed from their visible decay products and approximate independence
from uninteresting boosts, particularly those from the emission of QCD jets
outside of the supersymmetric decay tree.

Over the last several years, RJR event variables have been demonstrated as
effective tools for interpreting supersymmetric particle physics
data~\cite{santoni2018probing},
have been used in numerous \atlas\ analyses~\cite{%
atlas_rjr_SUSY_2016_07,
atlas_rjr_SUSY_2016_15,
atlas_rjr_SUSY_2016_16,
atlas_rjr_23l_SUSY_2017_03
}, and have influenced the designs of other
analyses~\cite{atlas_rjr_mimic_SUSY_2018_06}.

Decay trees used in the $\twoljets$ RJR analysis,
which displayed in Figure~\ref{fig:2ljets_rjr_decay_tres},
target C1N2-like scenarios similar to those also considered in the $\twoljets$
electroweak search.

>>>>

\begin{figure}[tp]
\centering
\begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figures/2ljets_sr_high_8_met_sig.pdf}
    \caption{SR-High-8}
\end{subfigure}
\hfill
\begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figures/2ljets_sr_llbb_mbb.pdf}
    \caption{SR-$\ell\ell bb$}
\end{subfigure}
\\[0.5em]
\begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figures/2ljets_sr_int_met_sig.pdf}
    \caption{SR-Int}
\end{subfigure}
\hfill
\begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figures/2ljets_sr_low_met_sig.pdf}
    \caption{SR-Low}
\end{subfigure}
\caption{%
Example signal region histograms with benchmark signal sample yields overlaid
as dotted lines~\cite{atlas2022searches}.
Physically, signal yields would add on top of the backgrounds.
Data are shown; regions in the the top two plots observe zero data, and
Poisson error bars for $0$ events are hidden for aesthetic reasons.
}
\label{fig:2ljets_signal_examples}
\end{figure}

@@

\begin{figure}[tp]
\centering
\includegraphics[width=0.99\textwidth]{figures/2ljets_contours_c1n2.pdf}
\caption{%
Contours for the C1N2 model in the $\twoljets$ electroweak
analysis~\cite{atlas2022searches}.
\\[0.5em]
Space below the solid red line is labelled as excluded and its dotted
neighbours show the result if all signal cross-sections are varied up and down
by theoretical error bars.
\\[0.5em]
The yellow band shows the $\pm1$-sigma region of exclusion contours
from asymptotic approximations to the prior distribution of the test statistic.
Grey areas are observed limits from the two-lepton parts
of~\cite{atlas_23l_SUSY_2016_24} and~\cite{atlas_2l_SUSY_2013_11}.
Exclusion is defined by the $95\%$ $\mathrm{CLs}$ prescription
in asymptotic approximations.
All contours are interpolated from a sparse grid.
}
\label{fig:2ljets_contours_c1n2}
\end{figure}

\begin{figure}[tp]
\centering
\includegraphics[width=0.99\textwidth]{figures/2ljets_contours_gmsb.pdf}
\caption{%
Contours for the GMSB model in the $\twoljets$ electroweak
analysis~\cite{atlas2022searches}.
\\[0.5em]
Space below the solid red line is labelled as excluded and its dotted
neighbours show the result if all signal cross-sections are varied up and down
by theoretical error bars.
\\[0.5em]
The yellow band shows the $\pm1$-sigma region of exclusion contours
from asymptotic approximations to the prior distribution of the test statistic.
Exclusion is defined by the $95\%$ $\mathrm{CLs}$ prescription
in asymptotic approximations.
All contours are interpolated from a sparse grid.
}
\label{fig:2ljets_contours_gmsb}
\end{figure}

@@

\clearpage
\FloatBarrier
\section{Origins}
% previous results on 2(/3)L from ATLAS
% other constraints on these models from ATLAS/CMS
% previous region design, motivation for this work
Previous analyses of \atlas\ data have explored similar selections of
two leptons, jets, and missing transverse momentum.
Quite a few analyses, in fact.
This project exists to supplement those results with information from the full
LHC Run 2 data-set which was completed in 2018 and contains
$139~\mathrm{fb}^{-1}$ of usable proton-proton collisions at $\sqrt{s} = 13~\eV[T]$.
Leverage of the increased data quantity is one key aim.
Analysis quality also might be improved if our designs can take lessons from
the experiences of previous work.

@@

In this way, RJR variables get some intelligence in how parent particles are
reconstructed from their visible decay products, as well as approximate
independence from uninteresting boosts, particularly those from the emission of
QCD jets outside of the supersymmetric decay tree.

Over the last several years, RJR event variables have been demonstrated to be
effective for interpreting supersymmetric particle physics
data~\cite{santoni2018probing},
and have been used in numerous \atlas\ analyses~\cite{%
atlas_rjr_SUSY_2016_07,
atlas_rjr_SUSY_2016_15,
atlas_rjr_SUSY_2016_16,
atlas_rjr_23l_SUSY_2017_03,
atlas_rjr_3l_SUSY_2019_09
} \TODO{more}.
They have also influenced the designs of other analyses which do not use
RJR variables themselves~\cite{atlas_rjr_mimic_SUSY_2018_06}.

Two decay trees are used in the $\twoljets$ RJR analysis, with signal region
each.
These trees target C1N2-like scenarios similar to those also considered in the
$\twoljets$ electroweak search,
and illustrated in Figure~\ref{fig:2ljets_rjr_decay_trees}.

\paragraph{Former excesses}
Excess data were observed in several signal regions of the partial Run 2
search ``for chargino--neutralino production using recursive
jigsaw reconstruction in final states with two or three charged
leptons''~\cite{atlas_rjr_23l_SUSY_2017_03},
which used $36.1~\mathrm{fb}^{-1}$ of proton-proton collisions at
$\sqrt{s} = 13~\eV[T]$.
That search reported notable excesses in four signal regions.
Of these, two required two charged leptons
($\mathrm{SR}2\ell\_\mathrm{Low}$ and $\mathrm{SR}2\ell\_\mathrm{ISR}$),
and other two required three
($\mathrm{SR}3\ell\_\mathrm{Low}$ and $\mathrm{SR}3\ell\_\mathrm{ISR}$).

Although the largest local excess was reported as ``$3.0$ standard deviations'',
which is not particularly large, the pattern of four related regions with
excess data can understandably draw attention.

Both $\mathrm{SR}3\ell\_\mathrm{Low}$ and $\mathrm{SR}3\ell\_\mathrm{ISR}$
were have been revisited with the full Run 2 data-set and updated background
modelling.
The update sees ``no significant excesses''~\cite{atlas_rjr_3l_SUSY_2019_09}.
These regions have also been approximated without direct use of RJR event
variables;
these approximations also find data ``in agreement'' with the background
model~\cite{atlas_rjr_mimic_SUSY_2018_06}.

The $\twoljets$ RJR serves to analysis reproduce
$\mathrm{SR}2\ell\_\mathrm{Low}$ and $\mathrm{SR}2\ell\_\mathrm{ISR}$,
with the full Run 2 data-set and updated background modelling,
and asks whether the excesses persist.
They do not.

\paragraph{RJR versus conventional}
Disappearing excesses have encouraged doubts about RJR variables;
the existence of the `RJR-mimic' paper~\cite{atlas_rjr_mimic_SUSY_2018_06}
is public evidence of this.

A more precise problem is that that phase space selected by RJR regions tends
to be poorly (inaccurately or imprecisely) modelled by simulations.
This is plausibly explained how RJR variables are not simply related to
conventional, lab-frame variables.
Simulations have been tuned to model conventional variables well, and
`systematic' variations also are evaluated and derived in conventional variables.
Not all functions of the data are well understood.

A lack of reliable simulations motivates reliance on `data-driven' background
estimates, which, from my perspective, are hard to interpret or trust.
For these reasons, the $\twoljets$ electroweak analysis uses only conventional
event variables.
